Redshift evolution of the SN stretch distribution

### 2019-12-19 Nicolas.tex
Big, BIG update

\usepackage{color}
\newcommand{\mr}[1]{{\textcolor[rgb]{0.60,0.10,0.6}{#1}}}
\newcommand{\mri}[1]{{\textcolor{red}{#1}}}

@@

@@

\abstract{Type Ia supernovae (SNe Ia) allow for the construction of the Hubble
    diagram, giving us information about the Universe's expansion and its
    fondamental components, one of which is dark energy. But systematic
    uncertainties are now starting to be limiting in our ability to measure
    those parameters. In particular, the physics of SNe Ia is still mostly
    unknown, and is thought not to change in time/with the redshift.}
    {In an attempt to reduce those uncertainties, we try to find an empirical
    law describing SNe Ia's length of explosion (stretch) evolution with the
redshift.}
    {We started by getting a complete sample representing all of the stretch
        distribution that Nature can give us, before using LsSFR measurments, an
        age tracer which evolution with redshift is known, that has been shown
        to have a strong correlation with the stretch. We compare their AICc, an
    estimator of the relative quality of statistical models that includes the
number of free parameters, to determine which ones describe besto the data.}
    {Models with an evolution of the stretch with the redshift have a better
    AICc than the ones without.}
    {We find that implementing these models allows us to fit the data better
    than models without stretch evolution.}

@@

%\begin{figure*}
%    \centering
%    \subfigure[Method]{\label{fig:zdet}
%        \includegraphics[width=8.5cm]{Article_figures/zmax_maglim_all.pdf}}
%    \subfigure[Result]{\label{fig:surveys_cuts}
%    \includegraphics[width=8.5cm]{A    \subfigure[Method]{\label{fig:zdet}
%rticle_figures/surveys_cuts_x.pdf}}
%    \subfigure[Method]{\label{fig:zdet}

@@

Type Ia supernovae (SNe Ia) are powerful cosmological distance indicators that
have enabled the discovery of the acceleration of the Universe's expansion
\citep{riess1998, permutter1999}. They remain today a key comological probe to
understand the properties of dark energy (DE) as it is the only tool able to
precisely map the recent expansion rate $z<0.5$, when DE is driving it
\citep[e.g.][]{scolnicastro2020}. They also are key to directly measure the
Hubble Constant (H$_0$) provided one can calibrate their absolute magnitude
\citep{riess2016, freedman2019}. Interestingly, the value of H$_0$ derived when
the SNeIa are anchored on Cepheids \citep[the SH0ES
project][]{riess2009,riess2016} is $\sim5~\sigma$ higher than what is predicted
by cosmic microwave background (CMB) data measured by Planck assuming the
standard $\L$CDM or when the SN luminosity is anchored at intermediate redshift
by the baryon accoustic oscillation (BAO) scale \citep{riess2019, ried2019,
planck, feeney}. While using the tip of the red giant branch technique in place
of the Cepheids seem to favor lower values of H$_0$ \citep{freedman,
recalibrationpaper} time delay measurements from strong lensing seem to also
favor high H$_0$ values \citep{wong2019}.

The H$_0$ tension has received a lot of attention as it could be a sign of new
fundamental physics. Yet, no simple solution is able to accomodate this H$_0$
tension when accounting for all other probes and the current most promising
scenario appears to be a burst of expansion at the matter-radiation decoupling
moment caused by a (fine tuned) early dark energy \citep{poulin}.

Alternatively, systematic effects affecting one or several of the aforementioned
analysis could also explain the tension. In \cite{rigault2015} we suggested that
SNe Ia from the cepheid calibrator sample differ significantly to the one from
the Hubble flow sample that enables to derive H$_0$. Indeed, Cepheid are very
young stars and the existence of Cepheids in a galaxy implies that the host is
star forming. In \citep{Rigault2015} we claimed that a more than 90\% of the
calibrator SNe Ia arise from a young-progenitor propulation while this
population only accounts for about half of the Hubble flow sample. Following
\cite{sullivan2010} and \citep{rigault2013}, we further showed that young and
old progenitor SNe Ia have a different magnitude of about 0.15 mag. This has
been confirmed for modern standardization technique
\citep[SALT2.4][]{guy2010,betoule2014} with high significance ($\sim 6\sigma$)
in \citep{rigault2018} as well as independently by \cite{roman2019} at $\sim
7\sigma$. As detailed in \citep{Rigault2015}, if the magnitude difference
between young and old SNe Ia is indeed of $0.15$ mag and if SNe Ia from the
calibrator sample is fully dominated by young SNeIa, the effect on H$_0$ of the
existance of these two SNe Ia populations is of 3.5\%. Accounting for the fact
that another environmental effect (the mass-step) is already included in the
SH0ES analysis, the net bias in the SH0ES analysis is expected closer to 3\%
\citep[see table 6 of ][]{rigault2015}.

SHOULD BE SHORTENED A LOT: However, the importance of this astrophysical bias in
the direct measurement of H$_0$ is still highly debated. We also highlight that
it could not explain the high value of H$_0$ derived by strong lensing. First,
\cite{jones2015} did not find any environmental bias when extending the
\citep{rigault2015} study. It was more recently reported in \citep{jones2019}
that indeed local host analysis has significant influence on the SN magnitude
but much more importantly than claimed in \citep{rigault2018} and with a
connection with the progenitor physics and/or a line of sight effect that is
unclear. More importantly, \cite{riess2016} attended to mimic the cepheid sample
selection function on the Hubble flow sample by removing SNe Ia from the latter
if they classificed their host as early-type or non-star forming. They measured
that doing so has no effect on the measurement of H$_0$.

% SERIES OF PAPER
This paper is part of a series of papers where we test the validity of the two
SNe Ia population models developped in \citep{rigault2018}.Briday et al. in prep
will present how using different environmental tracer is affecting our ability
to distinguish the two populations, and Rigault et al. in prep will further
study the connection with H$_0$. Here we analyze predictions concerning the
SNe Ia redshift drifts: (1) that the distribution of the SNIa stretch, a purely
intrinsic SNIa property, is evolving as a function of redshift and (2) that this
population-drift follows the relation that \citep{rigault2018} derived assuming
that one population is “young“ with a rate proportional to the star formation
rate in the Universe and the second is "old" with a rate proportional to the
stellar mass.

The concept of the SNeIa age dichotomy arised when the SNIa rate has first been
studied. \cite{Mannucci et al. 2005; Scannapieco & Bildsten 2005; Sullivan et
al. 2006; Aubourg et al. 2008} have shown that the relative SNe Ia rate in
galaxies could only be explained if two populations existed, one young,
following the host star formation activity, and one old following the host
stellar mass (the so called “prompt and delayed“ or “A+B“ model). In
\cite{rigault2018} we used the specific star formation rate at the SN location
(Local sSFR or LsSFR) to classify which are the prompts (those with a high
LsSFR) and which are the delayed (those with low LsSFR). Since the first SNe Ia
host analysis, the SN strech has been known to be strongly correlated with the
SN host properties \citep{Hamuy1996, Hamuy2000} and it has been extensively
confirmed since \citep[e.g.][]{Neill2009,Sullivan2010; Lampeitl2010; Kelly2010;
Gupta2011,D’Andrea2011; Childress2013,Rigault2013,Pan2014}. Following the “A+B“
model and the connection between SN stretch and host properties,
\citep{Howell20XX} first discussed the potential redshift drift of the SN strech
distribution. In this paper we revisit this analysis with the most recent SNe Ia
dataset and we test the validity of the two SNIa age populations model
developped by \citep{rigault2018}.

We present in section~\ref{sec:data} the sample we are using for this analysis,
which is based on the pantheon dataset \citep{scolnic2018a}. We discuss the
importance of obtaining a "complete" sample, i.e. reprensentative of the true
underlying SNe Ia distribution, and how we build one from the Pantheon sample.
We then present in section~\ref{sec:model} our modeling of the distribution of
stretch as a function of redshift based on the relative stretch distrubtion of
young and old SNeIa. Our results are presented in section~\ref{sec:result} where
we test if the distribution of stretch does evolve as a function of redshift and
if the age-model is in good agreement with this evolution. We discuss our
results and conclude in section~\ref{sec:conclusion}.

@@

\section{Sample}\label{sec:sam}

We base our analysis on the latest SNe Ia compilation: the Pantheon catalog from
\cite{scolnic2018}. Using this dataset, we aim at probing the evolution of the
distribution of SNe Ia stretch in nature as a function of redshift. \mr{We thus have to }

Because the
observed magnitude of a SN Ia is correlated with its stretch and its color, in a
magnitude limited survey such as those consituing most of the Pantheon sample,
the distribution of these lightcurve parameters is not representative of the
true underlying distribution above a certain limiting redshift. The first SNe Ia
that such a survey will miss because they will be too faint are indeed the
reddest and the lowest-stretch SNe Ia. If such a selection effect is not
accounted for, we might confuse true population drift we selection function or
conversally.

The figure~\ref{fig:maglim} shows the distribution of stretch for the magnitude
limited samples of the Pantheon data, i.e. PanStarrs \citep[PS1][]{ref}, the
Sloan Digital Sky Survey \citep[SDSS][]{ref} and the SuperNovae Legacy Survey
\citep[SNLS][]{astier2006}. An ellipse in the salt2.4 $(x_1, c)$ plan with major
axis $x_1=3$ and manor axis $c=0.3$ incapsulate the full distribution \citep[see
also][for a similar contour but using a less conservative
$c=0.2$]{campbell2013}. Assuming the average SN absolute magnitude ($x_1=0$,
$c=0$) is $M=-19.36$ \citep{kessler2009,scolnic2014} we can derive the absolute magnitude of
SNeIa along this ellipse given the standardisation coefficient from
\citep{scolnic2018a}: $\alpha=0.16$ and $\beta=3.14$.

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{Article_figures/zmax_maglim_all.pdf}
    \caption{SALT2.4 stretch and color lightcurve parameters of SNeIa from the
        SDSS, PS1 and SNLS samples from the pantheon catalog. The individual SN
        data are shown as blue dots and a 2D histogram is shown in grey to
        highlight point density. The ellipse encapsulating all the SN data is
        displayed colored by the effective standardized absolute magnitude using
    the $\alpha$ and $\beta$ standardisation coefficents from
\citep{scolnic2018}.}
    \label{fig:maglim}
\end{figure}

The faintest SNIa is that with $x_1=-1.66$ and $c=0.25$ and has an absolute \mr{standardized} magnitude at peak \mr{in Bessel-b band} of $-18.31$ mag, which leads to limiting asbolute magnitude of
$-18.00$ if we want to be able to observe it typically a week before and 10 days after peak.  
Therefore, given the $5\sigma$ point source detection magnitude
limit of a magnitude limited survey, one can derive the maximum redshift above which the faintest SNeIa will start to be missed.

\mr{SNLS typically acquires SNeIa in the redshift range $0.4<z<0.8$. At these redshifts the rest-frame Bessel-b band roughly corresponds to the SNLS-$i$ filter, that has a 24.8 mag $5\sigma$ depth \footnote{\href{https://www.cfht.hawaii.edu/Science/CFHTLS/cfhtlsfinalreleaseexecsummary.html}{CFHT final release website.}}. This converts to a $z_{lim}=0.60$, in perfect agreement with \citep{neill2006} and \citep{perrett2010}.
Similarly, PS1 observing SNeIa in the range $0.2<z<0.4$, their $g$-band $5\sigma$ depth is 23.1 mag \citep{rest2014}, corresponding $z_{lim}=0.30$.
The fig.~6 of \citep{scolnic2018} suggests a more conservative $z_{lim}$ of 0.27 for the PS1 catalog. In similar redshift range, SDSS has a limiting magnitude of 22.5 \citep{dilday2008,sake2008,sako2014}, which would lead to a $z_max=0.24$. However, the SDSS survey were more sensitive to limited spectroscopic ressources. \cite{kessler2009} presents that during year-1 of SDSS, SNeIa with $r-mag<20.5$ where favored for spectroscopic follow up, corresponding to the redshift cut at $0.15$. For the rest of the SDSS survey, additional spectroscopic ressources where used, such that \cite{kessler2009} and  \cite{dilday2008} of show a relative completness up to $z_{lim}=0.2$. For the rest of the analysis, we will use $z_{lim}=0.2$ as baseline magnitude.}

\mr{In Section~\ref{sec:testzmax}, we test the impact on using more conservative redshift limits ($z_{lim}=0.15$ for SDSS, $z_{lim}=0.27$ for PS and $z_{lim}=0.55$ for SNLS, following Fig.~14 of \citealt{perrett2010}).} The cuts on the raw data of those three samples are shown figure \red{fig:cuts}.

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{Article_figures/hist_surveys_cuts.pdf}
    \caption{Histograms of the 3 surveys on which we applied cuts to get rid of selection effects, as discussed in section \ref{sec:sam}.}
    \label{fig:cuts}
\end{figure}

In addition, we use the SNe Ia sample of from the nearby supernovae factory
\citep[SNfactory][]{aldering2004} published in \citep{rigault2018}. As the SNe
search were much deeper than the spectrophotometric follow-up, SNfactory SNe Ia
within a redshift range of $0.02<z<0.09$ should also be a random sampling of the
underlying SN population. The HST sample from
Pantheon follows the same logic and we therefore kept it entirely \citep{FIND BACK THE REF}.

Our "complete sample" is therefore made of 656 SNeIa whose composition is
detailed in table~\ref{tab:sample}.

\begin{table}
    \centering
    \caption{Origin of the SNeIa in our sample and redshift limit (if
    applicable) of each surveys.}
    \label{tab:sample}
    \begin{tabular}{l c c}
    \hline\hline\\[-0.8em]
        Survey & N$_{\mathrm{SN}}$  & z$_{lim}$ \\[0.15em]
        \hline\\[-0.8em]
        SNf & 141 & --\\[0.30em]
        SDSS & 167 & 0.20\\[0.30em]
        PS1 & 160 & 0.30 \\[0.30em]        
        SNLS & 102 & 0.60\\[0.30em]
        HST & 26 & --\\[0.30em]
        \hline
    \end{tabular}
\end{table}

%\mri{COMPARE HERE WITH LITERATURE. E.G. FIG. 6 OF SCOLNIC 2018, PS zmax AROUND0.27}

@@

%\mri{Dilday et al. 2008 | SDSS | sure good at z<0.12, fig. 10 suggest still okat z=0.2}

@@

\mri{CHECK LITERATURE CAN WE USE THE CfA SAMPLES ???}

\section{Method}

We used the LsSFR and stretch measurments from the SNf sample. The LsSFR has an
evolution with the redshift that is analytically known: calling $\de\z$ the
fraction of young stars and $\p\z$ the fraction of old ones, Rigault 2018 et +
find
\begin{equation}
    \frac{\delta\z}{\psi\z} \equiv \mathrm{LsSFR}\z = K\times \LF{1+z}^\varphi
\end{equation}
with $\f = 2.8$, and knowing $\de\z + \p\z = 1$:
\begin{align}
    \delta\z & = \left{K^{-1}\times \LF{1+z}^{-\f} +1\right}^{-1}\\
    \psi\z & = \LF{K\times \LF{1+z}^{+\f} + 1}\um
\end{align}

%\begin{figure*}
%    \centering
%    \subfigure[Raw data]{\label{fig:snf_data}
%        \includegraphics[width=8.5cm]{Article_figures/BiGaussian_onlydata.pdf}}
%    \subfigure[First implemented model]{\label{fig:3G2M2SSNF}
%        \includegraphics[width=8.5cm]{Article_figures/BiGaussian.pdf}}
%    \caption{Stretch des supernovae étudiées par la collaboration SNF en
%    fonction de log(LsSFR). La couleur représente la probabilité pour une
%supernova d'être issue d'un jeune progéniteur.}
%    \label{fig:mod_first}
%\end{figure*}

>>>>

\usepackage{color}
\newcommand{\mr}[1]{{\textcolor[rgb]{0.60,0.10,0.6}{#1}}}
\newcommand{\mri}[1]{{\textcolor{red}{#1}}}
\newcommand{\nn}[1]{{\textcolor[rgb]{1, 0.27, 0}{#1}}}

@@

\newcommand{\prob}[2]{\mathcal{P}\left( #1 \mid #2\right)}

@@

\abstract{Type Ia supernovae (SNe Ia) allow for the construction of the Hubble
    diagram, giving us information about the Universe's expansion and its
    fundamental components, one of which is dark energy. But systematic
    uncertainties are now starting to be limiting in our ability to measure
those parameters. In particular, the physics of SNe Ia is still mostly unknown,
and is thought not to change in time/with the redshift.}
    {In an attempt to reduce those uncertainties, we try to find an empirical
    law describing SNe Ia's length of explosion (stretch) evolution with the
redshift.}
    {We started by getting a complete sample representing all of the stretch
        distribution that Nature can give us, before using LsSFR measurements,
        an age tracer which evolution with redshift is known, that has been
        shown to have a strong correlation with the stretch. We compare their
        AICc, an estimator of the relative quality of statistical models that
    includes the number of free parameters, to determine which ones describe
best the data.}
    {Models with an evolution of the stretch with the redshift have a better
    AICc than the ones without.}
    {We find that implementing these models allows us to fit the data better
    than models without stretch evolution.}

@@

%\begin{figure*}
%    \centering
%    \subfigure[Method]{\label{fig:zdet}%        \includegraphics[width=8.5cm]{Article_figures/zmax_maglim_all.pdf}}

%        \includegraphics[width=8.5cm]{Article_figures/zmax_maglim_all.pdf}}
%    \subfigure[Result]{\label{fig:surveys_cuts}
%    \includegraphics[width=8.5cm]{A    \subfigure[Method]{\label{fig:zdet}
%rticle_figures/surveys_cuts_x.pdf}}
%    \subfigure[Method]{\label{fig:zdet}

@@

Type Ia supernovae (SNe Ia) are powerful cosmological distance indicators that
have enabled the discovery of the acceleration of the Universe's expansion
\citep{riess1998, perlmutter1999}. They remain today a key cosmological probe to
understand the properties of dark energy (DE) as it is the only tool able to
precisely map the recent expansion rate $z<0.5$, when DE is driving it
\citep[e.g.][]{scolnicastro2020}. They also are key to directly measure the
Hubble Constant (H$_0$) provided one can calibrate their absolute magnitude
\citep{riess2016, freedman2019}. Interestingly, the value of H$_0$ derived when
the SNeIa are anchored on Cepheids \citep[the SH0ES
project][]{riess2009,riess2016} is $\sim5~\sigma$ higher than what is predicted
by cosmic microwave background (CMB) data measured by Planck assuming the
standard $\L$CDM or when the SN luminosity is anchored at intermediate redshift
by the baryon acoustic oscillation (BAO) scale \citep{riess2019, reid2019,
planck2016, feeney} \nn{changed ried to reid; changed planck to planck2016;
feeney = https://tinyurl.com/vg6uufm ?}. While using the tip of the red giant
branch technique in place of the Cepheids seem to favor lower values of H$_0$
\citep{freedman2019, recalibrationpaper} \nn{changed "freedman" to
"freedman2019" ; recalibrationpaper = ?} time delay measurements from strong
lensing seem to also favor high H$_0$ values \citep{wong2019}.

The H$_0$ tension has received a lot of attention as it could be a sign of new
fundamental physics. Yet, no simple solution is able to accommodate this H$_0$
tension when accounting for all other probes and the current most promising
scenario appears to be a burst of expansion at the matter-radiation decoupling
moment caused by a (fine tuned) early dark energy \citep{poulin2019} \nn{changed
"poulin" to "poulin 2019"}.

Alternatively, systematic effects affecting one or several of the aforementioned
analysis could also explain the tension. In \cite{rigault2015} we suggested that
SNe Ia from the Cepheid calibrator sample differ significantly to the one from
the Hubble flow sample that enables to derive H$_0$. Indeed, Cepheid are very
young stars and the existence of Cepheids in a galaxy implies that the host is
star forming. In \citep{rigault2015} we claimed that a more than 90\% of the
calibrator SNe Ia arise from a young-progenitor population while this population
only accounts for about half of the Hubble flow sample. Following
\cite{sullivan2010} and \citep{rigault2013} \nn{should use cite or citep?}, we
further showed that young and old progenitor SNe Ia have a different magnitude
of about 0.15 mag. This has been confirmed for modern standardization technique
\citep[SALT2.4][]{guy2010,betoule2014} with high significance ($\sim 6\sigma$)
in \citep{rigault2018} as well as independently by \cite{roman2019} \nn{not
found on ADS: https://tinyurl.com/uemattg} at $\sim 7\sigma$. As detailed in
\citep{rigault2015}, if the magnitude difference between young and old SNe Ia is
indeed of $0.15$ mag and if SNe Ia from the calibrator sample is fully dominated
by young SNeIa, the effect on H$_0$ of the existence of these two SNe Ia
populations is of 3.5\%. Accounting for the fact that another environmental
effect (the mass-step) is already included in the SH0ES analysis, the net bias
in the SH0ES analysis is expected closer to 3\% \citep[see table 6
of][]{rigault2015}.

SHOULD BE SHORTENED A LOT: However, the importance of this astrophysical bias in
the direct measurement of H$_0$ is still highly debated. We also highlight that
it could not explain the high value of H$_0$ derived by strong lensing. First,
\cite{jones2015} \nn{put the ref but nos sure it's the correct one} did not find
any environmental bias when extending the \citep{rigault2015} study. It was more
recently reported in \citep{jones2019} \nn{put the ref but nos sure it's the
correct one} that indeed local host analysis has significant influence on the SN
magnitude but much more importantly than claimed in \citep{rigault2018} and with
a connection with the progenitor physics and/or a line of sight effect that is
unclear. More importantly, \cite{riess2016} attended to mimic the Cepheid sample
selection function on the Hubble flow sample by removing SNe Ia from the latter
if they classified their host as early-type or non-star forming. They measured
that doing so has no effect on the measurement of H$_0$.

This paper is part of a series of papers where we test the validity of the two
SNe Ia population models developed in \citep{rigault2018}.Briday et al. in prep
will present how using different environmental tracer is affecting our ability
to distinguish the two populations, and Rigault et al. in prep will further
study the connection with H$_0$. Here we analyze predictions concerning the SNe
Ia redshift drifts: (1) that the distribution of the SNIa stretch, a purely
intrinsic SNIa property, is evolving as a function of redshift and (2) that this
population-drift follows the relation that \citep{rigault2018} derived assuming
that one population is “young“ with a rate proportional to the star formation
rate in the Universe and the second is "old" with a rate proportional to the
stellar mass.

The concept of the SNeIa age dichotomy arised when the SNIa rate has first been
studied. \cite{mannucci2005, scannapieco2005, sullivan2006, aubourg2008} have
shown that the relative SNe Ia rate in galaxies could only be explained if two
populations existed, one young, following the host star formation activity, and
one old following the host stellar mass (the so called “prompt and delayed“ or
“A+B“ model). In \cite{rigault2018} we used the specific star formation rate at
the SN location (Local sSFR or LsSFR) to classify which are the prompts (those
with a high LsSFR) and which are the delayed (those with low LsSFR). Since the
first SNe Ia host analysis, the SN stretch has been known to be strongly
correlated with the SN host properties \citep{hamuy1996, hamuy2000} \nn{put
hamuy1996 as most cited of https://tinyurl.com/sjlbdz8} and it has been
extensively confirmed since \citep[e.g.][]{neill2009, sullivan2010,
lampeitl2010, kelly2010, gupta2011, dandrea2011, childress2013, rigault2013,
pan2014} \nn{added all the refs}. Following the “A+B“ model and the connection
between SN stretch and host properties, \citep{howell2009} \nn{changed
howell20xx to howell2009
(https://ui.adsabs.harvard.edu/abs/2009ApJ...691..661H/abstract)} first
discussed the potential redshift drift of the SN stretch distribution. In this
paper we revisit this analysis with the most recent SNe Ia dataset and we test
the validity of the two SNIa age populations model developed by
\citep{rigault2018}.

We present in section~\ref{sec:sample} the sample we are using for this
analysis, which is based on the pantheon dataset \citep{scolnic2018a}. We
discuss the importance of obtaining a "complete" sample, i.e. representative of
the true underlying SNe Ia distribution, and how we build one from the Pantheon
sample.  We then present in section~\ref{sec:modeling} our modeling of the
distribution of stretch as a function of redshift based on the relative stretch
distribution of young and old SNeIa. Our results are presented in
section~\ref{sec:results} where we test if the distribution of stretch does
evolve as a function of redshift and if the age-model is in good agreement with
this evolution. We discuss our results and conclude in section~\ref{sec:ccl}.

@@

\section{A "complete" Sample}
\label{sec:sample}

We base our analysis on the latest SNe Ia compilation: the Pantheon catalog from
\cite{scolnic2018a}.  \mr{TO BE REPHRASED} Using this dataset, we aim at probing
if the average stretch, an intrinsic SN Ia property, evolves as a function of
redshift or not; the underlying question being that of the SNIa population
redshift drift.  The stretch has been shown to depend on the SN environment star
formation activity
\citep[e.g.][]{howell2007,sullivan2010,childress2013,rigault2013,rigault2018}
and because the star formation rate is about 40 times larger at the high
redshift end of the pantheon sample than at its lower redshifts
\citep[e.g.,][and references therein]{tasca2015}, one might think that simply
comparing the redshifts distribution in a few bins of redshift could be
sufficient to probe any SN stretch redshift drift.  In practice, selection
effects are affecting the observed SN stretch distributions, which could
significantly differ from the true underlying one if these selection effects
become significant. Indeed, because the observed SNIa magnitude correlates with
the lightcurve stretches (and colors) the first SNIa that such a survey will
miss will the reddest and the lowest-stretch ones. Consequently, if selection
effects is not accounted for, one might confuse true population drift with
selection effect and conversely.

In a magnitude limited survey, providing sufficient (and random) spectroscopic
follow-up for acquiring typing and host redshift, the selection effects should
be negligible bellow a given redshift where all SNe Ia, even the faintest, could
be observed.  In contrast, targeted surveys have highly complex selection
functions and we will discard them from our analysis. Fortunately, modern SN
cosmology sample such has the pantheon dataset are now dominated by magnitude
limited survey.

We present fig.~\ref{fig:maglim} the lightcurve stretch and color distributions
of SNeIa from these surveys, namely: PanStarrs \citep[PS1][]{ref}, the Sloan
Digital Sky Survey \citep[SDSS][]{ref} and the SuperNovae Legacy Survey
\citep[SNLS][]{astier2006}.  An ellipse in the salt2.4 $(x_1, c)$ plan with
major axis $x_1=3$ and manor axis $c=0.3$ encapsulates the full distribution
(see also \citealt{bazin2011} and \citealt{campbell2013} for a similar contour
but using a less conservative $c=0.2$ for the latter). Assuming the SN absolute
magnitude with $x_1=0$ and  $c=0$ is $M=-19.36$ \citep{kessler2009,scolnic2014}
we can derive the absolute standardized magnitude at maximum of light along the
aforementioned ellipse given the standardization coefficient $\alpha=0.16$ and
$\beta=3.14$ from \cite{scolnic2018a}.  The faintest SNIa is that with
$x_1=-1.66$ and $c=0.25$ and has an absolute standardized magnitude at peak in
Bessel-b band of $-18.31\,\mathrm{mag}$. Since we need to see this object
typically a week before and 10 days after peak to build a suitable lightcurve,
the effective absolute limiting magnitude therefore is
$\sim-18.00\,\mathrm{mag}$.  Hence, given the $5\sigma$ point source detection
magnitude limit of a magnitude limited survey, one can derive the maximum
redshift above which the faintest SNeIa will start to be missed.

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{Article_figures/zmax_maglim_all.pdf}
    \caption{SALT2.4 stretch and color lightcurve parameters of SNeIa from the
        SDSS, PS1 and SNLS samples from the pantheon catalog. The individual SN
        data are shown as blue dots and a 2D histogram is shown in gray to
        highlight point density. The ellipse encapsulating all the SN data is
        displayed colored by the effective standardized absolute magnitude using
    the $\alpha$ and $\beta$ standardization coefficients from
\citep{scolnic2018a}.}
    \label{fig:maglim}
\end{figure}

SNLS typically acquires SNeIa in the redshift range $0.4<z<0.8$. At these
redshifts the rest-frame Bessel-b band roughly corresponds to the SNLS-$i$
filter that has a 24.8 mag $5\sigma$
depth\footnote{\href{https://www.cfht.hawaii.edu/Science/CFHTLS/cfhtlsfinalreleaseexecsummary.html}{CFHT
final release website.}}. This converts to a $z_{lim}=0.60$, in perfect
agreement with \cite{neill2006}, \cite{perrett2010} and \cite{bazin2011}.
Similarly, PS1 observes SNeIa in the range $0.2<z<0.4$, their $g$-band $5\sigma$
depth is 23.1 mag \citep{rest2014}, which yields to $z_{lim}=0.30$.  Figure~6 of
\citep{scolnic2018a} suggests a more conservative $z_{lim}$ of 0.27 for the PS1
catalog.  In similar redshift range, SDSS has a limiting magnitude of 22.5
\citep{dilday2008,sako2008}, which would lead to a $z_{lim}=0.24$. However, the
SDSS survey were more sensitive to limited spectroscopic resources.
\cite{kessler2009} present that during year-1 of SDSS, SNeIa with $r-mag<20.5$
were favored for spectroscopic follow up, corresponding to the redshift cut at
$0.15$. For the rest of the SDSS survey, additional spectroscopic resources were
used, such that \cite{kessler2009} and \cite{dilday2008} show a relative
completeness up to $z_{lim}=0.2$.  Following these analysis we will use
$z_{lim}=0.2$ as the baseline SDSS redshift limit for the rest of the analysis.
We could have chosen more conservative redshift limits for each of these
surveys, namely $z_{lim}=0.15$ for SDSS, $z_{lim}=0.27$ for PS and
$z_{lim}=0.55$ for SNLS, following Fig.~14 of \citealt{perrett2010}, and we will
also present results obtained when doing so.  We present the redshift
distribution of these three surveys in fig.~\ref{fig:cuts}, where we see that
the redshift limit roughly correspond to the pick of these histograms, as
expected.

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{Article_figures/hist_surveys_cuts.pdf}
    \caption{Histograms of the 3 surveys on which we applied cuts to get rid of
    selection effects, as discussed in section \ref{sec:sample}. The SNe Ia
between the conservative and non-conservative cuts are shown in transparent.}
    \label{fig:cuts}
\end{figure}

In addition, we use the SNeIa from the nearby supernovae factory
\citep[SNfactory][]{aldering2004} \nn{added ref assuming
https://ui.adsabs.harvard.edu/abs/2004APS..APR.V4003A/abstract} published in
\citep{rigault2018} and that have been discoverd from no-targeted survey (114
SNeIa, see their section~3 and 4.2.2). As the SNe search were much deeper than
the spectrophotometric follow-up, SNfactory SNe Ia within a redshift range of
$0.02<z<0.09$ should also be a random sampling of the underlying SN population.
Data from \citep{rigault2018} are within these redshift boundaries. The HST
sample from Pantheon follows the same logic and we therefore kept it entirely
\citep{FIND BACK THE REF} \nn{lots of choices: https://tinyurl.com/rcsx7cw}.

\begin{table}
    \centering
    \caption{Source surveys of SNeIa used in our "complete" sample. Conservative
    cuts are indicated in brackets.}
    \label{tab:sample}
    \begin{tabular}{l c c}
    \hline\hline\\[-0.8em]
        Survey & N$_{\mathrm{SN}}$  & z$_{max}$ \\[0.15em]
        \hline\\[-0.8em]
        SNf & 114 & --\\[0.30em]
        SDSS & 167(82) & 0.20(0.15)\\[0.30em]
        PS1 & 160(122) & 0.30(0.27) \\[0.30em]        
        SNLS & 102(78) & 0.60(0.55)\\[0.30em]
        HST & 26 & --\\[0.30em]
        \hline
    \end{tabular}
\end{table}

%%%%%%%%%%%%%%%%%%%%%%%
%   LITERATURE        %
%%%%%%%%%%%%%%%%%%%%%%%
%\mri{COMPARE HERE WITH LITERATURE. E.G. FIG. 6 OF SCOLNIC 2018, PS zmax
%AROUND0.27}

@@

%\mri{Dilday et al. 2008 | SDSS | sure good at z<0.12, fig. 10 suggest still
%okat z=0.2}

@@

%%%%%%%%%%%%%%%%%%%%%%%
%       INFO          %
%%%%%%%%%%%%%%%%%%%%%%%
%Fraction of random SDSS stretchs being < mean_conservative =  0.299
%Fraction of random PS1 stretchs being < mean_conservative =  0.254
%Fraction of random SNLS stretchs being < mean_conservative =  0.417

\section{Modeling the redshift drift}
\label{sec:modeling}

In \cite{rigault2018} we presented a modeling of the evolution of the fraction
of prompt and delayed SNeIa as a function of redshift following former work on
rates and delay time distributions; see e.g. the prompt and delayed SNeIa model,
aka the "A+B" model,
\cite{mannucci2005,scannapieco2005,sullivan2006,aubourg2008} and
\cite{maozmannucci2014} for a review on delay time distributions of SNeIa. In
short, we assume that the number of prompt SNeIa follows the star formation
activity in the Universe while the number of delayed SNeIa follows the number of
Gyr-old stars in the Universe, i.e. the stellar mass. Hence, if we denote
calling $\delta(z)$ (resp. $\phi(z)$) the fraction of young (resp. old) SNeIa in
the Universe as a function of redshift, then their ratio is expected to follow
the evolution of the specific star formation rate which goes as $(1+z)^{2.8}$
until $z\sim2$ \citep{tasca2015}. Hence since there is  $\sim50\%$ of young
SNeIa at $z\sim0.05$ \citep{rigault2013,rigault2018}, in agreement with rate
expectations \citep{mannucci2006,rodney2014}, then, we find in
\cite{rigault2018}:
\begin{align}
    \label{eq:delta}
    \delta(z) & = \left( K^{-1} \times (1+z)^{-2.8} +1 \right)^{-1}\,
                  \mathrm{or,}\\
    \psi(z) & = \left( K \times (1+z)^{+2.8} +1 \right)^{-1},
\end{align}

where $K=0.87$ given the fraction of young in the SNfactory sample. This
modelisation is comparable to the evolution predicted by \cite{childress2014}
based on SN rates in galaxies depending on their quenching time as a function of
their stellar mass.

\subsection{With LsSFR data}
\label{sec:modelpy}
